%&pdflatex

\RequirePackage{fix-cm}
% \RequirePackage{fixcmex}  % Commented out: not available in TeX Live
\documentclass{svjour3}                     %

\smartqed  %

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{hyperref}
\usepackage[numbers]{natbib}

\usepackage{orcidlink} 
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{argmax}
\usepackage{siunitx}

\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetKwRepeat{Do}{do}{while}
\usepackage{setspace} 
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{tcolorbox}
\begin{document}

\title{Detecting Multiple Semantic Concerns in Tangled Code Commits%
}


\author{
Beomsu Koh\orcidlink{0009-0004-2742-2200}
\and
Neil Walkinshaw\orcidlink{0000-0003-2134-6548}
\and
Donghwan Shin\orcidlink{0000-0002-0840-6449}
}


\institute{
Beomsu Koh \at
Department of Computer Science, University of Sheffield, Sheffield, United Kingdom \\
\email{gobeumsu@gmail.com}
\and
Neil Walkinshaw \at
Department of Computer Science, University of Sheffield, Sheffield, United Kingdom \\
\email{n.walkinshaw@sheffield.ac.uk}
\and
Donghwan Shin \at
Department of Computer Science, University of Sheffield, Sheffield, United Kingdom \\
\email{d.shin@sheffield.ac.uk}
}


\date{Received: date / Accepted: date}

\maketitle
\input{abstract}
\input{section1}
\input{section2}
\input{section3}
\input{section4}
\input{section5}
\input{section6}
\input{section7}






\section*{Declarations}

\subsection*{Funding}
The authors did not receive support from any organization for the submitted work.

\subsection*{Ethical Approval}
Not applicable.

\subsection*{Informed Consent}
Not applicable.

\subsection*{Author Contributions}
Beomsu Koh: Conceptualization, Methodology, Software, Formal analysis, Writing -- original draft.\\
Neil Walkinshaw: Supervision, Writing -- review \& editing.\\
Donghwan Shin: Supervision, Writing -- review \& editing.

\subsection*{Data Availability Statement}
All artefacts produced in this study, including the tangled commit dataset, fine-tuned models, and supporting scripts, have been made publicly available. See Section~\ref{sec:data-availability} for details.

\subsection*{Conflict of Interest}
The authors have no conflicts of interest to declare that are relevant to the content of this article.

\subsection*{Clinical Trial Number}
Not applicable.


\bibliographystyle{spbasic}      %
\bibliography{references} 


\end{document}